Exoplanet population inference \& the rate of Earth analogs from noisy, incomplete catalogs

### 2014-06-01 document/ms.tex
total refactor of abstract, small title tweaks

\title{%
    Exoplanet occurrence rate inference \& the rate density of Earth analogs
    from noisy incomplete catalogs
}
% \title{%
%   Inferring the distribution and rate of Earth-like exoplanets \\
%   from noisy individual-planet inferences
% }

\newcommand{\nyu}{2}
\newcommand{\mpia}{3}
\newcommand{\princeton}{4}
\newcommand{\berkeley}{5}
\author{%
    Daniel~Foreman-Mackey\altaffilmark{1,\nyu},
    David~W.~Hogg\altaffilmark{\nyu,\mpia},
    Timothy~D.~Morton\altaffilmark{\princeton}
}
\altaffiltext{1}         {To whom correspondence should be addressed:
                          \url{danfm@nyu.edu}}
\altaffiltext{\nyu}      {Center for Cosmology and Particle Physics,
                          Department of Physics, New York University,
                          4 Washington Place, New York, NY, 10003, USA}
\altaffiltext{\mpia}     {Max-Planck-Institut f\"ur Astronomie,
                          K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{\princeton}{Department of Astrophysics, Princeton University,
                          Princeton, NJ, 08544, USA}

@@

An impressive number of exoplanets have been discovered and it is of great
scientific interest to study what we can say about the exoplanet occurrence
rate function given these discoveries.
Every catalog is censored by non-trivial selection effects and detection
efficiencies, and the measurements of the system properties (period, radius,
\etc) are noisy.
We present a general hierarchical probabilistic framework for making justified
inferences about the occurrence rate of exoplanets, taking into account survey
completeness \emph{and observational uncertainties}.
By applying our method to synthetic catalogs, we demonstrate that it produces
more accurate estimates of the underlying occurrence rate than the standard
inverse-detection-efficiency weighting procedure.
When applied to an existing catalog of small planet candidates around G dwarfs
(Petigura \etal\ 2013), our method confirms previous measurements of a
flattening radius distribution near Earth's radius.
Although no true extra-solar Earth analog is known, hundreds of planets have
been found around Sun-like stars that are either Earth-sized but on shorter
periods, or else on year-long orbits but somewhat larger.
Under strong assumptions, these data have been used to make an extrapolated
estimate of the probability or rate at which Sun-like stars host Earth
analogs.
We relax some of these assumptions, only requiring that the occurrence rate be
a smooth function of period and radius (using a Gaussian process) and find
that the rate of Earth analogs per star is about 0.02 (per natural logarithmic
bin in period and radius) with large uncertainty.

>>>>

\title{%
    Exoplanet population inference \& the rate of Earth analogs
    from noisy, incomplete catalogs
}
% \title{%
%   Inferring the distribution and rate of Earth-like exoplanets \\
%   from noisy individual-planet inferences
% }

\newcommand{\nyu}{2}
\newcommand{\mpia}{3}
\newcommand{\cds}{4}
\newcommand{\princeton}{5}
\newcommand{\berkeley}{6}
\author{%
    Daniel~Foreman-Mackey\altaffilmark{1,\nyu},
    David~W.~Hogg\altaffilmark{\nyu,\mpia,\cds},
    Timothy~D.~Morton\altaffilmark{\princeton}
}
\altaffiltext{1}         {To whom correspondence should be addressed:
                          \url{danfm@nyu.edu}}
\altaffiltext{\nyu}      {Center for Cosmology and Particle Physics,
                          Department of Physics, New York University,
                          4 Washington Place, New York, NY, 10003, USA}
\altaffiltext{\mpia}     {Max-Planck-Institut f\"ur Astronomie,
                          K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{\cds}      {Center for Data Science,
                          New York University,
                          4 Washington Place, New York, NY, 10003, USA}
\altaffiltext{\princeton}{Department of Astrophysics, Princeton University,
                          Princeton, NJ, 08544, USA}

@@

Many exoplanets have been discovered and characterized;
it is now possible to study the population as a whole.
Although no true extra-Solar Earth analog is known, hundreds of planets have
been found around Sun-like stars that are either Earth-sized but on shorter
periods, or else on year-long orbits but somewhat larger.
Under strong assumptions, these exoplanet catalogs have been used to make an
extrapolated estimate of the probability or rate at which Sun-like stars host
Earth analogs.
Complicating these studies are that
every catalog is censored by non-trivial selection effects and detection
efficiencies, and every system property measurement (period, radius,
\etc) is noisy.
Here we present a general hierarchical probabilistic framework for making justified
inferences about the occurrence rate of exoplanets, taking into account survey
completeness and, for the first time, \emph{observational uncertainties}.
We are able to make fewer assumptions about the distribution than previous studies;
we only require that the occurrence rate be a smooth function of period
and radius (employing a Gaussian process).
By applying our method to synthetic catalogs, we demonstrate that it produces
more accurate estimates of the whole population than standard procedures based
on weighting by inverse detection efficiency.
We apply the method to an existing catalog of small planet candidates around
G dwarf stars (Petigura \etal\ 2013).
We confirm a previous result that the radius distribution changes slope near Earth radius.
We find
that the rate of Earth analogs is about 0.02 (per star per natural logarithmic
bin in period and radius) with large uncertainty.
This number is much smaller than previous estimates made with the same data but
stronger assumptions.
